---
title: "Reproducible Research: Peer Assessment 1"
author: "Juan Felipe Padilla Sepulveda"
date: "Fecha de actualización: `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    keep_md: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  dpi = 400,
  echo = TRUE,
  fig.align = "center",
  fig.asp = 0.618,
  fig.show = "hold",
  message = FALSE,
  out.width = "70%",
  tidy = "styler",
  warning = FALSE
)
```

```{r}
# Add packages
library(tidyverse)
library(skimr)
library(scales)
library(lubridate)

# Defaults
theme_set(hrbrthemes::theme_ipsum())
```

## Loading and preprocessing the data

```{r}
unzip("./activity.zip")

raw_data <- read_csv("./activity.csv")

skim(raw_data)
```

## What is mean total number of steps taken per day?

### Calculate the total number of steps taken per day

```{r}
days_data <- raw_data %>% 
    group_by(date) %>% 
    summarise(total_steps = sum(steps, na.rm = TRUE))
```

### Make a histogram of the total number of steps taken each day

```{r}
days_data %>% 
    ggplot(aes(total_steps)) +
        geom_histogram(bins = 10) +
        scale_x_continuous(labels = comma) +
        labs(x = "Steps by day", y = "Number of observations")
```

### Calculate and report the mean and median of the total number of steps taken per day

```{r}
days_data %>% 
    summarise(
        mean_steps = mean(total_steps, na.rm = TRUE),
        median_steps = median(total_steps, na.rm = TRUE)
    ) 
```     

## What is the average daily activity pattern?

### Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
raw_data %>% 
    group_by(interval) %>% 
    summarise(total_steps = mean(steps, na.rm = TRUE)) %>% 
    ggplot(aes(interval, total_steps)) +
        geom_line() +
        labs(x = "Interval", y = "Average number of steps taken")
```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
raw_data %>% 
    group_by(interval) %>% 
    summarise(total_steps = mean(steps, na.rm = TRUE)) %>% 
    arrange(desc(total_steps)) %>% 
    slice(1)
```

## Imputing missing values

### Calculate and report the total number of missing values in the dataset 

```{r}
raw_data %>% 
    filter(is.na(steps)) %>% 
    count(steps)
```

### Devise a strategy for filling in all of the missing values in the dataset

```{r}
imputed_data <- raw_data %>% 
    group_by(interval) %>% 
    mutate(mean_steps = mean(steps, na.rm = TRUE)) %>% 
    ungroup() %>% 
    mutate(
        steps = if_else(
            is.na(steps),
            mean_steps,
            steps
        )
    ) %>% 
    select(-mean_steps)
```

### Create a new dataset that is equal to the original dataset but with the missing data filled in

```{r}
imputed_data %>% 
    skim()
```

### Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day

```{r}
imputed_data %>% 
    group_by(date) %>% 
    summarise(total_steps = sum(steps, na.rm = TRUE)) %>% 
    ggplot(aes(total_steps)) +
        geom_histogram(bins = 10) +
        scale_x_continuous(labels = comma) +
        labs(x = "Steps by day", y = "Number of observations")
```

```{r}
imputed_data %>% 
    group_by(date) %>% 
    summarise(total_steps = sum(steps, na.rm = TRUE)) %>% 
    summarise(
        mean_steps = mean(total_steps, na.rm = TRUE),
        median_steps = median(total_steps, na.rm = TRUE)
    ) 
```

The values differ substantially from the previous calculations. Likewise, the distribution is now more normal.

## Are there differences in activity patterns between weekdays and weekends?

### Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day

```{r}
final_data <- imputed_data %>% 
    mutate(
        type_day = wday(date, label = TRUE, abbr = FALSE),
        type_day = if_else(
            type_day %in% c("sábado", "domingo"),
            "weekend",
            "weekday"
        ),
        type_day = factor(type_day, c("weekday", "weekend"))
    ) 
```

### Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r}
final_data %>% 
    group_by(interval, type_day) %>% 
    summarise(total_steps = mean(steps, na.rm = TRUE)) %>% 
    ggplot(aes(interval, total_steps)) +
        geom_line() +
        labs(x = "Interval", y = "Average number of steps taken") +
        facet_wrap(~ type_day, nrow = 2)
```
